### 2010-05-24 Documents/Publications/type-2/type-2.tex

\documentclass[12pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{times}
\usepackage[margin=0.75in]{geometry}

\title{A class of live and safe petri-nets that supports scalable
  concurrency analysis}

\author{Sameer D. Sahasrabuddhe \\ Kavi Arya \\ Madhav P. Desai}

\date{}

\begin{document}

\maketitle

\begin{abstract}
  \large

  Petri-nets are a compact way of representing concurrent sequences of
  events in a system, where events are represented as transitions in
  the corresponding petri-net. Given two events in a system, it is
  useful to determine whether they can occur simultaneously. But it is
  difficult to make such a determination in an arbitrary petri-net. We
  present a class of petri-nets called ``Type-2'' which restricts the
  structure of the petri-net in order to enable a scalable analysis,
  without severely affecting the expressive power of the petri-net.
  Given a Type-2 petri-net, it is possible to determine in linear time
  whether two transitions can fire simultaneously. In addition, it is
  also possible to generate an exhaustive listing of sets of
  transitions such that no two transitions in each set can fire
  simultaneously.

  We first define the concept of a ``token-preserving region'' which
  produces a live and safe petri-net when augmented by a marked place.
  We further define a set of production rules to combine such regions
  into larger token-preserving regions. The petri-net created by
  augmenting such a region is said to be of the class ``Type-2''. The
  initial marking of the petri-net is restricted to a single marked
  place.

  The restriction on the structure of the Type-2 petri-net allow a
  simple labelling scheme to effectively simulates an execution of the
  petri-net. These labels can be compared in linear time to determine
  the concurrency of the labelled transitions. Despite the
  restrictions imposed by the production rules, the Type-2 class of
  petri-nets is powerful enough to represent synchronous systems. For
  example, this class has been used as a transition step in
  translating imperative programs to hardware descriptions.

\end{abstract}

\end{document}