### 2010-05-24 Documents/Reports/Kunal-Sreenivas/conclusion.tex

\clearpage

\section{Conclusion}

From the results, tabulated and plotted earlier, we can conclude the following:
\begin{enumerate}
\item The nonpipelined memory looks more efficient in power-delay and area-delay, when compared to the deeply pipelined case. 
There's no significant improvement in delays as we increase the number of banks except A5/1.
This indicates that there is limited parallelism in the memory accesses.
\item The circuits are usually more energy efficient at higher frequency of operation.
\item The floating point operators limit the higher operating frequency in the case of FFT and Linpack.
\item Energy per operation for FFT and Linpack is in the order of 10 nJ and 12 nJ respectively. 
\end{enumerate}