### 2010-05-24 Documents/Reports/Kunal-Sreenivas/old_tite.tex

\documentclass[12pt]{report}      % Comments after  % are ignored

\usepackage{amsmath,amssymb,amsfonts} % Typical maths resource packages

\usepackage{graphics}                 % Packages to allow inclusion of graphics

\usepackage{color}                    % For creating coloured text and background

\usepackage{hyperref}                 % For creating hyperlinks in cross references

\oddsidemargin 0cm

\evensidemargin 0cm

\pagestyle{myheadings}         % Option to put page headers

                               % Needed \documentclass[a4paper,twoside]{article}

\markboth{{\small\it PDF\ \LaTeX \ coloured text and graphics}}

{{\small\it C.T.J. Dodson} }

\textwidth 15.5cm

\topmargin -1cm

\parindent 0cm

\textheight 24cm

\parskip 1mm

\newtheorem{theorem}{Theorem}[section]

\newtheorem{proposition}[theorem]{Proposition}

\newtheorem{corollary}[theorem]{Corollary}

\newtheorem{lemma}[theorem]{Lemma}

\newtheorem{remark}[theorem]{Remark}

\newtheorem{definition}[theorem]{Definition}

\def\R{\mathbb{ R}}

\def\S{\mathbb{ S}}

\date{\small\it May 16, 2000}

\title{\fcolorbox{white}{white}{\color{black}Power / Delay Estimation of Auto Generated Circuits}}

%\footnote{A demonstration example including colored text and graphics}}}

\author{ K. P. Ghosh, S.Subramanian \\

{\small Department of Electrical Engineering, IIT, Bombay, India}

 }