### 2010-05-24 Documents/Reports/Kunal-Sreenivas/ref.tex

\addcontentsline{toc}{chapter}{References}
\begin{thebibliography}{99}

\bibitem{} Premkishore Shivakumar, Norman P. Jouppi. Cacti 3.0 : An Integreted Cache Timing, Power, and Area model,
{\bf Western Research Laboratory}, 250 University Avenue Palo Alto, California 94301 USA.

\bibitem{} Banit Agrawal, Timothy Sherwood. Guiding Architectural SRAM models, 
{\bf Published in International Conference of Computer Design (ICCD) 2006},Department of Computer Science, University of California, Santa Barbara.

\end{thebibliography}