### 2010-05-24 Documents/Reports/Kunal-Sreenivas/report.tex

\documentclass[conference, onecolumn]{IEEEtran}     % Comments after  % are ignored
\usepackage{amsmath,amssymb,amsfonts} % Typical maths resource packages
\usepackage{graphics}                 % Packages to allow inclusion of graphics
\usepackage{color}                    % For creating coloured text and background
\usepackage{hyperref}                 % For creating hyperlinks in cross references

\oddsidemargin 0cm
\evensidemargin 0cm

\pagestyle{myheadings}         % Option to put headers
                               % Needed \documentclass[a4paper,twoside]{article}

%\markboth{{\small\it PDF\ \LaTeX \ coloured text and graphics}}

%{{\small\it C.T.J. Dodson} }
\textwidth 15.5cm
\topmargin -1cm
\parindent 0cm
\textheight 24cm
\parskip 1mm

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}[theorem]{Definition}

\def\R{\mathbb{ R}}

\def\S{\mathbb{ S}}

\date{\small\it April 20, 2009}

\title{{Power / Delay Estimation of \\ Auto Generated Circuits}}

%\footnote{A demonstration example including colored text and graphics}}}

\author{ K. P. Ghosh, S.Subramanian \\
{\small Department of Electrical Engineering, IIT, Bombay, India}
 }

\begin{document}

\pagenumbering{arabic}
\maketitle
\begin{abstract}

The goal of the research project was to characterize RTL generated by the AHIR flow.
We have taken a set of examples, used AHIR to generate RTL, mapped the RTL to layout and finally performed power, area and delay estimation for the above examples.
For all our circuit examples we have used the OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler to perform synthesis.
The netlist thus obtained is used as an input to Cadence SOC Encounter to generate the layout.
The layout being closest to the actual hardware is most suiatble for estimation of power consumed in the design for a particular input sequence.
Five examples viz. A5, AES, Red Black trees, Linpack and FFT were selected, six memory architectures were explored for each and results are summarized in final section.

\end{abstract}
\input{introduction.tex}
\input{memory.tex}
\input{arch_choice.tex}
\input{a5.tex}
\input{linpack.tex}
\input{rb_trees.tex}
\input{fft.tex}
\input{aes.tex}
\input{conclusion.tex}
\input{ref.tex}
\end{document}