### 2010-07-13 Documents/Publications/type-2/type-2.tex
type-2: Introduction and early description of Type-1 petri nets

\documentclass[12pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{times}
\usepackage{amsfonts}
\usepackage[margin=0.75in]{geometry}

\title{On the use of petri nets for modelling high-level programs}

\author{Sameer D. Sahasrabuddhe \\ Kavi Arya \\ Madhav P. Desai}

\date{}

\newtheorem{definition}{Definition}

\newtheorem{srule}{Rule}

\begin{document}

\maketitle

\begin{abstract}

  Petri nets provide a powerful mechanism for the modelling of
  concurrent systems. However, because of their generality, it is
  often difficult to prove properties about petri net models unless
  some restrictions are placed on their construction. We consider the
  use of petri nets to model algorithms described in high level
  programming languages. In particular, we introduce a class of live
  and safe petri nets (which we term as {\em Type-2} petri nets) for
  this purpose. This class is defined by a set of production rules
  which ensure liveness and safety of the resulting petri net.

  Further, we show that Type-2 petri nets support an efficient method
  for concurrency analysis that uses a labelling scheme which is
  analogous to a symbolic simulation of the petri net. Thus, it is
  easy to determine the sets of events (or transitions) which satisfy
  ordering relationships (two transitions are said to be ordered if it
  is impossible for both of them to fire simultaneously). This is
  useful in identifying possibilities for conflict free resource
  sharing when implementing the programs being modelled.

  Despite the restrictions imposed by the production rules, the Type-2
  class of petri nets is powerful enough to represent programs written
  in traditional imperative languages such as C/C++ and also in
  synchronous languages such as Esterel. It has been used as a
  transition step in translating imperative programs to hardware
  descriptions\cite{ahir-thesis}. The resulting high-level synthesis
  flow uses the labelling scheme to optimise the generated hardware by
  identifying opportunities for sharing hardware resources in the
  absence of timing information.

\end{abstract}

\section{Introduction}
\label{sec:introduction}

Petri nets provide a powerful formalism that has been used to model
and analyse the behaviour of a large class of systems. We consider the
use of petri nets to model the behaviour of algorithms described in
high-level programming languages in order to derive efficient
implementations. We use the petri net to capture the control flow in
the program which is coupled with a data-path (a set of computational
resources) in order to implement the intended behaviour. The petri net
model can be analysed to unveil properties that are important when
deriving an implementation, such as safety and boundedness. In
addition, the analysis can also provide information for efficient
utilisation of resources in the data-path.

However, the complexity of analysing a petri net in general quickly
increases with its size. We adopt the strategy of reducing this
complexity by restricting the structure of the petri net itself. We
propose a class of petri nets (that we call ``Type-2'') that is live
and safe, and can be easily analysed to determine concurrency
relationships between different computations in the program.

The structure of the Type-2 petri net is well suited for representing
the control flow observed in high-level programs --- in particular,
the interaction between concurrency and choice (branches). Threads
that are started within a branch must synchronise before exiting the
branch, and conversely, the targets of a branch encountered in a
thread must be restricted to that thread. A Type-2 petri net is built
using a set of construction rules that restrict the structure of the
petri net to ensure this behaviour. The same restrictions enable an
efficient labelling scheme that can be used to analyse the concurrency
of operations in the program. Note that the behaviour allowed by a
Type-2 petri net is a generalised version of a series-parallel
net\cite{something}, as can be seen in Section~\cite{sec:somewhere}.

\section{Definitions}
\label{sec:defintions}

A petri net is a directed bipartite graph with two kinds of nodes ---
\textit{places} and \textit{transitions}. A \textit{marking} is a
function that assigns a non-negative number to each place, indicating
the number of tokens in that place.

\begin{eqnarray*}
  N & = & (P, T, E, M)\\
  \textrm{where } P & = & \textrm{finite set of places}\\
  T & = & \textrm{finite set of transitions}\\
  E & \subseteq & (P \times T) \cup (T \times P)\\
  M & : & P \rightarrow \mathbb{N}
\end{eqnarray*}

The above defintion allows at most one edge from a place (transition)
to a transition (place). A \textit{simple place (transition)} is a
place (transition) with one incoming edge and one outgoing place. A
\textit{branch (fork)} is a place (transition) with one incoming edge
and multiple outgoing edges, while a \textit{merge (join)} is a place
(transition) with multiple incoming edges and one outgoing edge.

\section{Type-1 petri nets}
\label{sec:type-1}

We first define the class of Type-1 petri nets, which are constructed
by a set of construction rules combined with the most basic Type-1
petri net. The class of Type-2 petri net is a subclass that is defined
by an additional construction rule introduced in
Section~\ref{sec:type-2}.

\begin{definition}
  A \emph{simple cycle} is the petri net $N = (P, T, E, M_0)$ where $P
  = \{p\}$, $T =\{t\}$, $E = \{(p,t), (t,p)\}$ and $M_0(p) = 1$.
\end{definition}

Clearly, a simple cycle is live and safe; and it is defined as the
simplest type-1 petri net. A type-1 petri net is defined as the petri
net obtained by repeatedly applying one more construction rules to the
simple cycle. Each rule replaces a simple place or transition in a
type-1 petri net with a petri net subgraph called a ``standard
region''. There are three types of standard regions, defined below.
The places occuring in a standard region are \textbf{not marked} in
the initial marking.

\begin{description}
\item [Series region:]
  A \emph{p-type series region} (t-type series region) is a petri net
  consiting of a sequence of simple transitions and places, where the
  first and the last elements are both places (transitions) with only
  one edge within the region. The first and the last elements
  are of the series region are called the \emph{entry} and the
  \emph{exit} of the region respectively.

\item [Branch region:]
  A \emph{branch region} is a conected petri net where every
  transition is simple, while every place is either a branch or a
  merge. No place in the branch region has an edge incident on an
  element outside the region, except the \emph{entry} which has one
  incoming edge from outside the region and the \emph{exit} which has
  one outgoing edge outside the region.

\item [Fork region:]
  A \emph{fork region} is a connected petri net where every place is
  simple, while every transition is either a fork or a join. No
  transition in a fork region has an edge incident on an element
  outside the region, execep the \emph{entry} and \emph{exit}
  transitions, analogous to the entry and exit of a branch region. In
  addition, a fork region is required to be a DAG.
\end{description}

The construction rules of a Type-1 petri net can now be stated as the
following replacements:

\begin{srule}
  The petri net obtained by replacing a simple place (transition) in a
  Type-1 petri net by a p-type series region (t-type series region) is
  a also a Type-1 petri net.
\end{srule}

\begin{srule}
  The petri net obtained by replacing a simple place in a Type-1 petri
  net by a branch region is also a Type-1 petri net.
\end{srule}

\begin{srule}
  The petri net obtained by replacing a simple transition in a Type-1
  petri net by a fork region is also a Type-1 petri net.
\end{srule}

\section{Liveness and Safety in a Type-1 petri net}
\label{sec:live-safe}

We prove that a Type-1 petri net is live and safe by induction over
the structure of the petri net. In particular, we show that the
replacement of a simple transition or place in a Type-1 petri net by
applying any of the construction rules does not affect the liveness
and safety of the rest of the petri net.

\subsection{Series Region}
\label{sec:series-live-and-safe}

\subsection{Branch Region}
\label{sec:branch-live-and-safe}

A petri net formed by a branch region connected to simple marked place
is a \emph{state machine} which is strongly connected and has exactly
one token. Hence it is live and safe\ref{murata-tutorial}.

\subsection{Fork Region}
\label{sec:fork-live-and-safe}

All cycles in a petri net formed by a fork region connected to a
simple marked place $p_0$ pass through the place $p_0$. Thus every
cycle in the petri net has a marked place (hence it is live) and every
place is on a cycle with exactly one marked place (hence it is safe).

\end{document}