### 2013-03-01 Documents/Publications/dsd2013/paper.tex
Added examples for dsd2013 paper..

\documentclass[conference]{IEEEtran} 
\usepackage{graphicx}
\usepackage[tight,footnotesize]{subfigure}
\usepackage{amsmath}
\usepackage{alltt}
\usepackage{multirow}
\usepackage{times}
\usepackage{url}

\title{On the impact of loop optimizations in the mapping of high-level programs to hardware}

\author{Madhav P. Desai\\
  Indian Institute of Technology -- Bombay, Powai, Mumbai -- 400076, INDIA\\
  Email: madhav@ee.iitb.ac.in }
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]

\newcommand{\sym}[1]{$\operatorname{#1}$}

\newbox\subfigbox             % Create a box to hold the subfigure.
\makeatletter
  \newenvironment{subfloat}% % Create the new environment.
    {\def\caption##1{\gdef\subcapsave{\relax##1}}%
     \let\subcapsave=\@empty % Save the subcaption text.
     \let\sf@oldlabel=\label
     \def\label##1{\xdef\sublabsave{\noexpand\label{##1}}}%
     \let\sublabsave\relax    % Save the label key.
     \setbox\subfigbox\hbox
       \bgroup}%              % Open the box...
      {\egroup                % ... close the box and call \subfigure.
     \let\label=\sf@oldlabel
     \subfigure[\subcapsave]{\box\subfigbox\sublabsave}}%
\makeatother

\pagestyle{empty}

\begin{document}

\maketitle
\thispagestyle{empty}

\begin{abstract}

  In the context of mapping high-level algorithms to hardware,
  we consider the basic problem of generating an efficient hardware 
  implementation of a loop.  We describe a control-flow mechanism through
  which multiple iterations of a loop can be dynamically pipelined in
  the generated hardware,  and study the extent to which this 
  mechanism affects the properties of the generated hardware.
  First, we observe that the additional resources needed by this
  mechanism are small.  Further, we observe that this mechanism, when
  combined with static loop unrolling, offers an order-of-magnitude improvement 
  in performance for common kernels such as the fast-fourier transform,
  matrix multiplication and inner product computations.  
  Finally, we observe that these loop-optimization techniques in
  yield efficient mappings of single threads to hardware.  %% need more stuff here.

\end{abstract}

\section{Introduction}

%% The importance and problem of loops.
A high-level-synthesis (HLS) compiler takes an algorithm described in
a high-level programming language (such as C, for example) and produces a circuit
description (a register-transfer-level description in a hardware
description language such as VHDL, for example).
High-level synthesis systems have been available for some time
now (for example, \cite{pegasus-cash}, \cite{spark-vlsi-paper} and
commercial tools such as Catapult-C\footnote{\url{http://www.mentor.com/products/esl/high_level_synthesis/}} from Mentor Graphics
C-to-Silicon\footnote{\url{http://www.cadence.com/products/sd/silicon_compiler/}} from Cadence,
Synopsys HLS\footnote{\url{http://www.synopsys.com/Tools/SLD/HLS/}} from Synopsys).

The main attraction of such a compiler is a substantial reduction in the design and verification effort.  The
main concern is the quality of the hardware that is produced.
For the same starting point, the quality of the hardware produced may be objectively
measured in terms of factors such as area, energy-consumption and performance (in the
sense of time required for completion, or in terms of the rate at which the hardware
finishes the computational tasks).  An {\em internal measure} of the quality of the
hardware produced is the utilization factor of the major resources in the hardware (that is,
the utilization factor of expensive resources).

%% How compilers have handled this problem.
In this paper, we consider the problem of mapping loops to hardware.  It
is well known that most compute intensive programs spend a large fraction
of their time in inner loops.  Thus, the optimization of these loops is
a primary problem for any optimizing compiler which is targetting
either a deeply pipelined or parallel processor.  Several loop optimizations
have been considered in literature, such as loop-unrolling, loop-peeling,
software-loop-pipelining etc. \cite{wolfe}, \cite{muchnick}, with the intent of
these optimizations being the extraction of as much
parallelism as possible from the source program.

%% Algorithm-to-hardware:  can similar techniques work?
One expects that any technique which extracts parallelism from
the source program would also be useful in improving the efficiency
of hardware generated from that source program.  We consider two
such loop optimizations:
\begin{itemize}
\item {\em Loop-unrolling}, a {\em static} compile-time optimization technique:  
an inner loop is unrolled by instantiating multiple copies of the loop-body
while simultaneously reducing the number of loop-iterations.  For example:
\begin{verbatim}
for(i=0; i < 8; i++) {
   x += a[i]*b[i];
}
\end{verbatim}
is transformed to
\begin{verbatim}
for(i=0; i < 4; i+=2) {
   x += a[i]*b[i];
   x += a[i+1]*b[i+1];
}
\end{verbatim}
This unrolling increases the size of the basic block (that is, the maximal sequence of
statements without any branches), and provides the possibility of extracting more
parallelism in the loop.
\item {Loop-pipelining}, a {dynamic} run-time optimization in hardware:  the hardware
itself starts executing multiple iterations of the loop in parallel, while respecting
all dependencies within and across loop iterations.   This is analogous to software
loop-pipelining, except that software-loop-pipelining is static in nature, and is
determined at compile time, whereas we are looking at a dynamic technique in which
the generated hardware itself keeps multiple loop iterations in flight while maintaining
dependencies within and across loop iterations.
\end{itemize}

In the remainder of this paper, we will first briefly describe the model of
the hardware that is produced by our HLS compiler and illustrate how this model
can incorporate dynamic run-time loop pipelining.  The chief issues here are the
hardware overhead (area, energy, delay) incurred by the need to provide 
this run-time support in the hardware generated by the compiler, and
the corresponding improvement in performance that results from this optimization.

In order to address this issue, we will present a set of observations from
experiments performed on representative inner loops that occur in 
some important applications such as the fast-fourier transform, the
matrix product, vector dot-product, and a digital filtering algorithm.
These observations report the hardware performance on four different
loop-optimization choices: with no loop optimization, with static unrolling
alone, with dynamic loop pipelining alone, and with static unrolling combined with
dynamic loop pipelining.  Hardware resource utilization and delays are
computed by synthesizing the generated hardware for an FPGA target.

The observations indicate the following:
\begin{itemize}
\item The performance improvement with loop-pipelining alone is in the
2X-3X range, with a small increase in the hardware resources needed.
\item The performance improvement with loop-unrolling alone is in the 
2X range, but substantial hardware overheads are incurred.
\item The combination of loop-pipelining and loop-unrolling leads
to a performance improvement between 8-10X, with the hardware overheads
being similar to those incurred with unrolling alone.
\end{itemize}

%%
%% TODO..  so what?
%%

\section{Related work}

%% compiler techniques.

%% hls techniques to handle loops?

\section{Model of hardware generated by our compiler}

The hardware generated by our compiler consists of three
cooperating components: the control-path, the data-path and
the storage system \cite{ahir}

To illustrate the model, we consider a simple example.
\begin{verbatim}
float a[1024], b[1024];
float dotp = 0.0;
for(i=0; i < 1024; i++)
{
   dotp += a[i]*b[i];
}
\end{verbatim}
To compile this code, we use the clang-2.8 \cite{clang} C compiler, 
which is used to emit LLVM \cite{llvm} byte-code.  The LLVM byte-code
is transformed through a series of steps by our compiler tools to
produce the final hardware, which is depicted in Figure \ref{fig:dotp}.

%%
%%  control path and data-flow
%%

\subsection{Control-path}

%% 
%% Petri-net, constructed according to
%% certain rules.  Req/Ack transitions
%%

\subsection{Data-path}

%%
%% Directed hyper-graph.  Nodes are operators
%% hyper-arcs are wires.  Each operator node
%% has a split protocol.
%%

\subsection{Memory-subsystem}

%%
%% time-stamping scheme, in order completion.
%%

\subsection{An example of a loop}

%%
%% for(i=0;i < 100; i++)
%% {
%%    c[i] = a[i] + b[i];
%% }
%%

\section{Control-flow mechanism for loop-pipelining}

%%
%% reenable rules.
%%
%% loop-terminator.
%%
%% advantages of dynamic pipelining: any loop
%% can be handled..
%%

\section{Experimental Results}

%%
%% examples: vector add, dot-product, matrix-multiply, linfinity-norm.
%%
%%
%%       show results without-unrolling-pipelining, with-unrolling-pipelining, etc.
%%

\section{Conclusion}

%%
%% loop-pipelining and loop-unrolling have high impact as
%% in pipelined processors.
%%
%% hardware mechanism for loop-pipelining is cheap and effective.
%%
%% can get 10X improvement and very high utilization of 
%% hardware resources.
%%

\bibliography{ref}
\bibliographystyle{IEEEtran}

\end{document}

% LocalWords:  Req Ack LL DP CP cmpgt br cdfg cp dp ir intra req init NCA SA mW
% LocalWords:  TPR STPR STPRs ccccc RTL ASIC petri ack versa liveness LRG LLVM
% LocalWords:  VHDL SRAM Synopsys AES LPK Linpack FFT RBT Synposys OSU TSMC nJ
% LocalWords:  nm